Towards a classification framework for social machines

### 2013-02-26 latex/paper_new.tex
another intermediary version

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{img/socialmachinescope.png}
\caption{Social machines and related areas} \label{fig:social machine}
\end{center}
\end{figure}

@@

In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
'For Science', and $5$ one that mainly serves  'socializing' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

While determining the intersection of elements was straightforward, the consolidation of the
constructs required a more thorough process. We manually grouped the constructs into rough clusters, based around the areas they cover. We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to cover the same aspects. The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set. This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters:

\begin{description}
  \item [Popularity]: this cluster mainly covered descriptive information about a given system, such as its maturity and number of users. For the purpose of our analysis we decided to discard the associated constructs, as they 
  \item [Tasks and purpose]
  \item [Motivation and incentives]
  \item [Technology]
\end{description}

The paper presents the current version of the framework. We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences. We provide initial evidence of the usefulness of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will

\subsection{The Polyarchical Relationship of Social Machines}

While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
social machines to be created, and instantiations of those frameworks into social machines that have been created for a specific purpose, such as Wikipedia
and Ushahidi Haiti. However, through further investigation, we discovered that the distinction between framework and instantiations was not the only
differentiator. Specifically, that some social machines could also have sub-machines or communities within them, that are working to solve different problems,
within the context of the larger machine. For example, Wikipedia is an instantiation of the MediaWiki framework, but there are a large number of
communities within Wikipedia that work on domain-specific areas, such as those contributing in areas of science and technology\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}. However, we must also note that wikipedia is primarily split into communities of language, with the English, German and French wikipedias having the largest number of articles. It is therefore possible to analyse wikipedia into a hierarchy, from MediaWiki, to the languages of Wikipedia, down to the domain-specific communities. We explored this idea further, to include other social machines, and links where social machines utilise other social machines in order to function. For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.  We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi platforms) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8.5cm]{img/polyarchy.pdf}
\caption{A polyarchy of Social Machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
\end{center}
\end{figure}

This approach enables us to start with a more detailed analysis of certain levels over others;
and seeing what similarities flow up and down the polyarchy. For example, what do
specific instances of Ushahidi/Zooniverse/MediaWiki have in common with other instances, and
how to they differ dynamically?

One originally unforeseen classification is that of the causes that utilise Social Machines.
Through a particularly popular cause, large numbers of participants can be mobilised, for
either a short period of time, or for long-term engagement, using multiple social machines
in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
self-archive their own papers online so they are available for free, in addition to hosted
versions behind publishers pay-walls. In order to achieve this aim, the social machine of
``open access'' typically uses the social machines of institutional repositories provided
by the authors' associated institutions. As with other social machines, these social
machines typically use off-the-shelf implementations (in this case the most popular are
EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
web. This cause is mature, growing year-on-year and has global coverage.

%\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
%
%%We identified a hierarchy of constructs pertaining to the motivation people have for using social
%%machines with top level constructs of:
%%
%%\begin{enumerate}
%%\item Hedonistic
%%\item Financial
%%\item To be informed
%%\item To help others
%%\end{enumerate}
%%
%%The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
%%
%%\begin{figure*}[htb]
%%\begin{center}
%%\includegraphics[width=18cm]{img/motivation.pdf}
%%\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
%%\end{center}
%%\end{figure*}
%
%The motivations of users can be identified and mapped to existing human needs and motivations
%systems, such as Maslow's hierarchy of human needs~\cite{maslow} (e.g., self-actualisation, security
%respect of others). For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
%
%\begin{enumerate}
%\item Benefit to contributor
%\item Benefit to moderator
%\item Benefit to system operator/host (e.g, Google, Facebook, Amazon)
%\item To affiliates of the host (e.g., Amazon affiliates, YouTube partners)
%\item To society as a whole
%\item To a contributor's social network
%\end{enumerate}
%
%In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
%we suggest that there are also resulting actions from each motivation. Thus, it should be
%possible to follow basic human needs to motivations in social machines, through to the
%actual actions (and therefore the necessary software feature support) that are made by
%participants in social machines. Our theory is that this path may enable designers of
%social machines to satisfy human needs by implemented specific features into their social
%machines, or to identify potential gaps in their serving of needs and motivations due
%to lack of implemented features.
%
%Motivation in specific social machines has been studied, including motivation of participants
%in: Wikipedia~\cite{kuznetsov2006motivations}; Open Source Software~\cite{lakhani2003hackers};
%Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
%Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
%
%[relate this work back to ours briefly.]

@@

\subsection{Constructs}

The clustering process described in Section~\ref{meth} yielded both a
reduction in the overall number of constructs (in particular an
elimination of overrepresented areas) but also identificaiton of four
simple themes: sets of constructs pertaining to users and roles within
each machine, second, the nature of tasks, problems and activities
performed through and using the machine, third, motivation for
participation, finally, the contexts of participation.  Each of these
clusters contain 5-8 constructs as illustrated in Table
\ref{table:constructs}; we describe each of these themes in the
following subsections.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{8cm}|}
\hline
{\bf Users and Roles} \\
\hline
Generality of Audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organisation of roles \\
Clear separation of roles/responsibilities among participants \\
\hline
{\bf Tasks} \\
\hline
Activities involve creative production of content \\
Activities involve subjective appraisal of content \\
Activities involve solving (a definable) computation task or set of problems\\
Tasks are domain-specific \\
The machine owner (service/platform) derives benefit from participation \\
Activities and tasks are pre-defined or participant-defined \\
Variation in types of contributions and tasks \\
\hline
{\bf Motivation} \\
\hline
Participants are intrinsically motivated:\\
\hspace{1cm} to gain/share knowledge \\
\hspace{1cm} to ``get something done'' \\
\hspace{1cm} to ``for fun/entertainment''\\
\hspace{1cm} to ``be social'' \\
\hspace{1cm} for the benefit of a specific group of people who need help \\
\hspace{1cm} for the benefit of society as a whole \\
Participation is motivated by extrinsic reward (payment, status etc) \\
\hline
{\bf Context(s) of Participation} \\
\hline
Participants' participate to fulfill needs of a role  \\
\hspace{1cm} work-related/professional role \\
\hspace{1cm} home/family related \\
\hspace{1cm} social role \\
\hspace{1cm} leisure/entertainment role \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hspace{1cm} Physical location is relevant to the service \\
\hline
\end{tabular}
\end{center}
\caption{Consolidated constructs of social machines.} \label{table:constructs}
\end{table}

\subsubsection{Users and Roles}
The first set of constructs pertain to the human participants of
social machines and the way(s) participants are organised within them.
One construct identifies whether participants constitute members of
the general public, or are of a specific demongraphic, external group
or organisation, occupation, or of a particular expertise.  Two
constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
to which interaction is constrained among participants, and to the
degree to which participants' identities are used and disclosed within
the machine. The final two constructs pertain to the roles: first,
as to whether multiple roles are pre-differentiated by the machine or
whether all participants essentially initially assume the same role,
second whether roles (pre-defined or emergent) are archical.

\subsubsection{Tasks, Problems and Activities}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, andf so on.  The resulting computation, looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Futher, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture these
difeferences, we include a construct that differentiates machines with
pre-defined problem spaces or are open to participant-provided tasks
and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

\subsubsection{Motivation}
The third set of themes pertains to the motivation that sustains
continued participation in these systems.  We identiy 6 different
kinds of common sources of intrinsic motivation; first, that
participation is fun, second that it accomplishes an activity that the
participant enjoys or wanted to get done, third, that it satisfies the
desire to gain or share knowledge, or, fourth, or to be social.  The
last two pertain to philanthropy, whether participation is seen to
benefit a particular group of disadvantaged people or individuals who
need assistance, or, finally, if participation is beneficial to
society as a whole.  Finally, we ask whether extrinsic motivational
factors also contributed to sustained participation - such factors
could include money reward, status, recognition and so on

\subsubsection{Context of Participation}
The final set of constructs pertains to the context(s) of each
participant's greater (life-) activities that participation is
performed.  This context could be in a work context, home/family
context, for leisure, and at a desk or away from a computer (in a
mobile context).  PArticipation could be conducted via apps, web
browsers, sensors and other technolgoies.  Due to the predominance of
``geo-locative'' services that incorporate information about the
user's location, we have included a construct identifying the degree
to which the user's location is used by the machine.

>>>>

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.3\textwidth]{img/socialmachinescope.png}
\caption{Social machines and related areas} \label{fig:social machine}
\end{center}
\end{figure}

@@

In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
'For Science', and $5$ one that mainly serves  'socializing' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

While determining the intersection of elements was straightforward, the consolidation of the
constructs required a more thorough process. We manually grouped the constructs into rough clusters, based around the areas they cover. We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to cover the same aspects. The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set. This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives. In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users. Further on, the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavors) in the grids provided by the ten experts, others such as the different types of workflows synthesizing human and automatized capabilities of a social machine were less taken into account. As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other. Extending the framework with constructs capturing these aspects (see, for instance, \cite{quinn2011human} for examples in the context of human computation) is part of our future work.

The paper presents the current version of the framework. We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences. We provide initial evidence of the usability of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will determine the extent to which the classification constructs identified can be meaningfully used by researchers and system designers and operators. We will test the completeness, correctness, and comprehensibility of the constructs in experiments in which a new set of social machines will be classified by framework users; we will ask the participants to assess the quality of the framework along these general dimensions, and measure inter-annotator agreement to learn about the usefulness of the classifications produced.

\subsection{The polyarchical relationship of social machines}
When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP ('games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators. A second observation is concerned with the relationship between broader and narrower-scoped instances of social machines; general-purpose examples such as Facebook, Twitter or Amazon, or even the Web, enable the formation of more specific social machines and communities within them, that are working to solve solve different problems. An example illustrating this relationship is the Obama'12 US Presidential campaign~\cite{obamakieron}, which relied on Twitter and Facebook as basic social machines; most forms of customer relationship management and community engagement carried out in the digital space use similar tools. Finally, social machines are interconnected into a greater ecosystem, both at the social and at the computation levels. For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia. In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualization of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.  We propose looking at nested machines with 'The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as infrastructure, frameworks, services, and projects and initiatives.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8.5cm]{img/polyarchy.pdf}
\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
\end{center}
\end{figure}

This approach enables us to start with a more detailed analysis of certain levels over others;
and seeing what similarities flow up and down the polyarchy. For example, what do
specific instances of Ushahidi/Zooniverse/MediaWiki have in common with other instances, and
how to they differ dynamically? How do certain design decisions taken at the level of the infrastructure, frameworks and service propagate into narrower-focused systems that are built on top of them? Similarly, how will such decisions affect a broader ecosystem of social machines, each with their own, though overlapping, purposes and communities?

The polyarchy is complementary to the classification constructs we elicitated. The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels. For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivize participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia. In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.

%One such example is the 'open access' movement in scholarly publishing
%~\cite{harnad2001self}. This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls. In order to achieve this aim, the social machine of 'open access' typically uses the social machines of institutional repositories provided
%by the authors' associated institutions. As with other social machines, these social
%machines typically use off-the-shelf implementations (in this case the most popular are
%EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
%web. This cause is mature, growing year-on-year and has global coverage.

%One originally unforeseen classification is that of the causes that utilise Social Machines.
%Through a particularly popular cause, large numbers of participants can be mobilised, for
%either a short period of time, or for long-term engagement, using multiple social machines
%in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
%at the lowest level of the polyarchy. 
%\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
%
%%We identified a hierarchy of constructs pertaining to the motivation people have for using social
%%machines with top level constructs of:
%%
%%\begin{enumerate}
%%\item Hedonistic
%%\item Financial
%%\item To be informed
%%\item To help others
%%\end{enumerate}
%%
%%The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
%%
%%\begin{figure*}[htb]
%%\begin{center}
%%\includegraphics[width=18cm]{img/motivation.pdf}
%%\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
%%\end{center}
%%\end{figure*}
%
%The motivations of users can be identified and mapped to existing human needs and motivations
%systems, such as Maslow's hierarchy of human needs~\cite{maslow} (e.g., self-actualisation, security
%respect of others). For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
%
%\begin{enumerate}
%\item Benefit to contributor
%\item Benefit to moderator
%\item Benefit to system operator/host (e.g, Google, Facebook, Amazon)
%\item To affiliates of the host (e.g., Amazon affiliates, YouTube partners)
%\item To society as a whole
%\item To a contributor's social network
%\end{enumerate}
%
%In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
%we suggest that there are also resulting actions from each motivation. Thus, it should be
%possible to follow basic human needs to motivations in social machines, through to the
%actual actions (and therefore the necessary software feature support) that are made by
%participants in social machines. Our theory is that this path may enable designers of
%social machines to satisfy human needs by implemented specific features into their social
%machines, or to identify potential gaps in their serving of needs and motivations due
%to lack of implemented features.
%
%Motivation in specific social machines has been studied, including motivation of participants
%in: Wikipedia~\cite{kuznetsov2006motivations}; Open Source Software~\cite{lakhani2003hackers};
%Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
%Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
%
%[relate this work back to ours briefly.]

@@

\subsection{Constructs}
The clustering process described in Section~\ref{meth} yielded both a
reduction in the overall number of constructs (in particular an
elimination of over-represented areas) and the identification of four
simple themes: sets of constructs pertaining to users and roles within
each machine, second, the nature of tasks, problems and activities
performed through and using the machine, third, motivation for
participation, finally, the contexts of participation.

Each of these clusters contain between five and eight constructs as illustrated in Table
\ref{table:constructs}; we describe each of these themes in the
following subsections.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{8cm}|}
\hline

{\bf Tasks, purpose and context of participation} \\
\hline
Activities involve creative production of content \\
Activities involve subjective appraisal of content \\
Activities involve solving (a definable) computation task or set of problems\\
Tasks are domain-specific \\
The machine owner (service/platform) derives benefit from participation \\
Activities and tasks are pre-defined or participant-defined \\
Variation in types of contributions and tasks \\
Participants' participate to fulfill needs of a role  \\
\hspace{1cm} work-related/professional role \\
\hspace{1cm} home/family related \\
\hspace{1cm} social role \\
\hspace{1cm} leisure/entertainment role \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hspace{1cm} Physical location is relevant to the service \\
\hline
{\bf Participants, roles and incentives} \\
\hline
Generality of audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organization of roles \\
Clear separation of roles/responsibilities among participants \\
Participants are intrinsically motivated:\\
\hspace{1cm} to gain/share knowledge \\
\hspace{1cm} to ``get something done'' \\
\hspace{1cm} to ``for fun/entertainment''\\
\hspace{1cm} to ``be social'' \\
\hspace{1cm} for the benefit of a specific group of people who need help \\
\hspace{1cm} for the benefit of society as a whole \\
Participation is motivated by extrinsic reward (payment, status etc) \\
\hline

\end{tabular}
\end{center}
\caption{Consolidated constructs of social machines.} \label{table:constructs}
\end{table}

\subsubsection{System view: tasks, purpose and context of participation}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, andf so on.  The resulting computation, looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Futher, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture these
difeferences, we include a construct that differentiates machines with
pre-defined problem spaces or are open to participant-provided tasks
and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

The final set of constructs pertains to the context(s) of each
participant's greater (life-) activities that participation is
performed.  This context could be in a work context, home/family
context, for leisure, and at a desk or away from a computer (in a
mobile context).  PArticipation could be conducted via apps, web
browsers, sensors and other technolgoies.  Due to the predominance of
``geo-locative'' services that incorporate information about the
user's location, we have included a construct identifying the degree
to which the user's location is used by the machine.

\subsubsection{Social view: users, roles, motivation and incentives}

The first set of constructs pertain to the human participants of
social machines and the way(s) participants are organised within them.
One construct identifies whether participants constitute members of
the general public, or are of a specific demongraphic, external group
or organisation, occupation, or of a particular expertise.  Two
constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
to which interaction is constrained among participants, and to the
degree to which participants' identities are used and disclosed within
the machine. The final two constructs pertain to the roles: first,
as to whether multiple roles are pre-differentiated by the machine or
whether all participants essentially initially assume the same role,
second whether roles (pre-defined or emergent) are archical.

The third set of themes pertains to the motivation that sustains
continued participation in these systems.  We identiy 6 different
kinds of common sources of intrinsic motivation; first, that
participation is fun, second that it accomplishes an activity that the
participant enjoys or wanted to get done, third, that it satisfies the
desire to gain or share knowledge, or, fourth, or to be social.  The
last two pertain to philanthropy, whether participation is seen to
benefit a particular group of disadvantaged people or individuals who
need assistance, or, finally, if participation is beneficial to
society as a whole.  Finally, we ask whether extrinsic motivational
factors also contributed to sustained participation - such factors
could include money reward, status, recognition and so on
